---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abcoxp

<!-- badges: start -->
<!-- badges: end -->

The goal of abcoxp is to implement approximate Bayesian inference for Cox Proportional Hazard model on partial likelihood function, that accommodates linear fixed effect, nonlinear covariate effect and random frailty.

## Installation

You can install the development version of this package from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("AgueroZZ/abcoxp")
```
## Example

This is a basic example which shows you how to use `abcoxp` to fit and analyze some models, we consider the following data set of kidney infection times from the R package `survival`:

```{r example, message=FALSE}
library(abcoxp)
library(dplyr)
## basic example code
data <- survival::kidney
head(data)
```


We can fit a simple fixed effect model with:

```{r fixed-model}
mod <- abcoxp_fit(data = data, 
                  times = "time", # name of your time variable in the data
                  cens = "status", # name of your censoring indicator (0/1) in the data
                  fixed = c("age", "sex"), # names of fixed effect (have to be numeric variables)
                  fixed_control = list(betaprec = 0.001) # the prior distribution for the fixed effects
                  )
```


Once the model is fitted, we can do Bayesian inference by sampling from its (approximate) posterior with:
```{r}
samps <- abcoxp_sampling(fitted_model = mod, # output from abcoxp_fit
                         nsamps = 3000 # number of independent posterior samples to draw
                         )
par(mfrow = c(1,2))
hist(samps$fixed_samps[,1], main = "Samples for effect of age", xlab = "Effect")
hist(samps$fixed_samps[,2], main = "Samples for effect of sex", xlab = "Effect")
```


One can also consider adding a group-level frailty and a nonlinear covariate effect with continuous second order random walk (RW2) covariance with:

```{r full-model, message=FALSE}
mod <- abcoxp_fit(data = data, 
                  times = "time", # name of your time variable in the data
                  cens = "status", # name of your censoring indicator (0/1) in the data
                  fixed = c("sex"), # names of fixed effect (have to be numeric variables)
                  fixed_control = list(betaprec = 0.001), # the prior distribution for the fixed effects
                  frailty = "id", # name of the group variable for frailty, has to be integer valued
                  RW2 = "age", # name of the RW2 variable, has to be numeric valued
                  frailty_control = list(u = 1, alpha = 0.5), # PC prior parameters for frailty SD
                  RW2_control = list(u = 1, alpha = 0.5, r = 30), # PC prior paramters for RW2 SD, and # of knots
                  Inference_control = list(aghq_k = 4) # number of grid points to use for the AGHQ
                  )
```


The posterior for the two hyperparameters (theta1 for SD of frailty and theta2 for SD of RW2) can be plotted by:

```{r plot1}
par(mfrow = c(1,4))
plot(mod$model)
```


Similarly the posterior samples can be obtained with:

```{r}
samps <- abcoxp_sampling(fitted_model = mod, # output from abcoxp_fit
                         nsamps = 3000 # number of independent posterior samples to draw
                         )
```

The variable `samps$RW2_values_samps` has posterior sample paths for the nonlinear covariate effect:

```{r}
pos_samp_path <- samps$RW2_values_samps
mean <- pos_samp_path[,-1] %>% apply(1, mean) # The first column is the covariate values, not sample path!
upper <- pos_samp_path[,-1] %>% apply(1, quantile, p = 0.975)
lower <- pos_samp_path[,-1] %>% apply(1, quantile, p = 0.025)
plot(mean ~ pos_samp_path$x, type = 'l', ylim = c(-1,1), xlab = "age", ylab = "Effect")
lines(upper ~ pos_samp_path$x, lty = 'dashed')
lines(lower ~ pos_samp_path$x, lty = 'dashed')

```










